Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning \& Rotation

### 2020-12-28 xrb2/paper.tex
Update affiliation
Incorporate some suggestions from last meeting

\author[0000-0001-6191-4285]{Kiran Eiden}
\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
             Stony Brook, NY 11794-3800}

@@

\begin{deluxetable}{lrrr}
	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each simulation. $s_0$ is the speed at $t = 0$. The initial flame velocities and accelerations are derived from a quadratic least-squares fit to each of the datasets for times $t \gtrsim 0.02~\mathrm{s}$.}
	\tablehead{\colhead{run} & \colhead{$s_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)}} 
	\startdata
	{\tt F1000}     & $(3.38 \pm 0.11) \times 10^5$ & $(-0.43 \pm 1.57) \times 10^5$ \\
	{\tt F500}      & $(2.66 \pm 0.27) \times 10^5$ & $(10.97 \pm 3.22) \times 10^5$ \\
	{\tt F500\_2E8} & $(3.98 \pm 0.19) \times 10^5$ & $(-1.56 \pm 2.63) \times 10^5$ \\
	{\tt F500\_3E8} & $(-4.56 \pm 2.92) \times 10^5$ & $(3.39 \pm 0.60) \times 10^7$ \\
	\enddata
	%
\end{deluxetable}

\begin{figure}[t]
	\centering
	\plotone{flame_speeds_rotation.pdf}
	\caption{\label{fig:flame_speeds_1} Flame front position vs.\ time for the standard ($10^8~\mathrm{K}$) 500 Hz and 1000 Hz runs. The dashed lines
		show quadratic least-squares fits to the data for $t \gtrsim 0.02~\mathrm{s}$.}
\end{figure}

Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{F500}$ and $\tt{F1000}$ runs (blue and orange, respectively) to show the dependence on rotation rate. As some of the hotter 500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data for times $t \gtrsim 0.02~\mathrm{s}$ and calculate the acceleration and initial velocity from the fit parameters. These results are given in Table~\ref{table:flame_speeds}. \MarginPar{How is the acceleration calculated for this table? I.e. Specifically, at what time / over what time window is it calculated?} \MarginPar{NF: I will update the error estimates}.

We found in \citet{flame_wave1} that the flame speed $s$ obeys:
\begin{equation}
	\label{eqn:rossby}
	s \appropto L_R (\epsdot)^{\half} \propto \frac{(\epsdot)^{\half}}{\Omega}
\end{equation}
where $\epsdot$ is the specific energy generation rate and $L_R$ is the Rossby length. This was consistent with the results of \citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby length. As seen in Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling of the flame speed with rotation rate in this set of runs. We observed earlier, however, that nuclear reactions progress more quickly at the higher rotation rate. This results in a higher $\enucdot$ --- up to four to six times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the increased Coriolis confinement. The initial position of $\tt{F1000}$ does appear to be reduced compared to the lower rotation $\tt{F500}$ run. From Equation \ref{eqn:rossby}, increasing the rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the initial position of the flame should be reduced by a similar factor (again, due to greater confinement from the Coriolis force). $L_R$ is not, however, a perfect metric for the flame lengthscale. The scaling described in Equation \ref{eqn:rossby} seems to be more accurate at higher rotation rates (1000 Hz and up), based on \citet{flame_wave1}. We also observe that $\tt{F500}$ accelerates faster than $\tt{F1000}$, which appears to experience a small deceleration at early times. \MarginPar{This is speculation on my part -- does someone else have clear idea why?} This disparity may be a direct result of the difference in Coriolis force. It could also be due to the 500 Hz flame taking more time to develop fully and reach higher burning rates, resulting in an $\epsdot$ \MarginPar{Is this supposed to be enucdot???}increase over time.

@@

\begin{figure}[t]
	\centering
	\plotone{500Hz_stacked_enuc1_2order.pdf}
	\caption{\label{fig:flame_speeds_2} Flame front position vs.\ time for the three 500 Hz runs with
		$T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data
		for $t \gtrsim 0.02~\mathrm{s}$. Note that, due to its rapid acceleration, $\tt{F500\_3E8}$ (orange) is only run out to $\sim 0.07~\mathrm{s}$ to avoid surpassing the domain boundary.}
\end{figure}

>>>>

\author[0000-0001-6191-4285]{Kiran Eiden}
\affiliation{Dept.\ of Astronomy, University of California, Berkeley,
	CA 94720-3411}
\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
	Stony Brook, NY 11794-3800}

@@

\begin{deluxetable}{lrrr}
	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each 
	simulation. $s_0$ is the speed at $t = 0$. The initial flame velocities and accelerations are 
	derived from a quadratic least-squares fit to each of the datasets for times $t \gtrsim 
	20~\mathrm{ms}$.}
	\tablehead{\colhead{run} & \colhead{$s_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)}} 
	\startdata
	{\tt F1000}     & $(3.38 \pm 0.11) \times 10^5$ & $(-0.43 \pm 1.57) \times 10^5$ \\
	{\tt F500}      & $(2.66 \pm 0.27) \times 10^5$ & $(10.97 \pm 3.22) \times 10^5$ \\
	{\tt F500\_2E8} & $(3.98 \pm 0.19) \times 10^5$ & $(-1.56 \pm 2.63) \times 10^5$ \\
	{\tt F500\_3E8} & $(-4.56 \pm 2.92) \times 10^5$ & $(3.39 \pm 0.60) \times 10^7$ \\
	\enddata
	%
\end{deluxetable}

\begin{figure}[t]
	\centering
	\plotone{flame_speeds_rotation.pdf}
	\caption{\label{fig:flame_speeds_1} Flame front position vs.\ time for the standard ($10^8~\mathrm{K}$) 500 Hz and 1000 Hz runs. The dashed lines
		show quadratic least-squares fits to the data for $t \gtrsim 20~\mathrm{ms}$.}
\end{figure}

Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{F500}$ and $\tt{F1000}$ runs 
(blue and orange, respectively) to show the dependence on rotation rate. As some of the hotter 
500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data for 
times $t \gtrsim 20~\mathrm{ms}$, with the fit parameters giving the flame acceleration and 
initial velocity. \MarginPar{KE: Can add an explanation of the averaging method once that's 
decided.} These results are displayed in Table~\ref{table:flame_speeds}. \MarginPar{NF: I will 
update the error estimates.}

We found in \citet{flame_wave1} that the flame speed $s$ obeys:
\begin{equation}
	\label{eqn:flame_speed_scaling}
	s \appropto L_R (\enucdot)^{\half} \propto \frac{(\enucdot)^{\half}}{\Omega}
\end{equation}
where $\enucdot$ is the specific nuclear energy generation rate and $L_R$ is the Rossby length. 
This was consistent with the results of \citet{cavecchi:2013}, who noted that the flame speeds in 
their simulations scaled with the Rossby length. As seen in Figure~\ref{fig:flame_speeds_1}, there 
is no obvious inverse scaling of the flame speed with rotation rate in this set of runs. We 
observed earlier, however, that nuclear reactions progress more quickly at the higher rotation 
rate. This results in a higher $\enucdot$ --- up to four to six times higher near the burning front 
after the flame ignites --- which may counteract the reduction in flame speed from the increased 
Coriolis confinement. The initial position of $\tt{F1000}$ does appear to be reduced compared to 
the lower rotation $\tt{F500}$ run. From Equation \ref{eqn:flame_speed_scaling}, increasing the 
rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the initial 
position of the flame should be reduced by a similar factor (again, due to greater confinement from 
the Coriolis force). $L_R$ is not, however, a perfect metric for the flame lengthscale. The scaling 
described in Equation \ref{eqn:flame_speed_scaling} seems to be more accurate at higher rotation 
rates (1000 Hz and up), based on \citet{flame_wave1}. We also observe that $\tt{F500}$ accelerates 
faster than $\tt{F1000}$, which appears to experience a small deceleration at early times. 
\MarginPar{KE: This is speculation on my part -- does someone else have clear idea why?} This 
disparity may be a direct result of the difference in Coriolis force. It could also be due to the 
500 Hz flame taking more time to develop fully and reach higher burning rates, resulting in an 
$\enucdot$ increase over time.

@@

\begin{figure}[t]
	\centering
	\plotone{500Hz_stacked_enuc1_2order.pdf}
	\caption{\label{fig:flame_speeds_2} Flame front position vs.\ time for the three 500 Hz runs with
		$T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data
		for $t \gtrsim 20~\mathrm{ms}$. Note that due to its rapid acceleration, $\tt{F500\_3E8}$ 
		(orange) is only run out to $\sim 0.07~\mathrm{s}$ to avoid surpassing the domain boundary.}
\end{figure}
